---
title: "Binomial, Negative Binomial Practice Problems"
format: 
  html:
    self-contained: true
editor_options: 
  chunk_output_type: console
knitr: 
  opts_chunk: 
    dev: svg
---

1. The probability of completing a difficult PhD program is 0.7.  Calculate the probability that in a cohort of 29 students, 3 or fewer students do not complete the degree.

```{r}

```

1. In this difficult program, students complete the degree after 5 years (no variation).  A professor decides to stop taking new students after advising 10 students to completion.  She also decides to retire after all of her active students graduate.  If the professor only takes 1 student a year (at the start of the program), what is the probability that the professor will be retired in 20 years?
```{r}

```


1. 3 prizes are awarded to the first 3 correct solutions of a difficult problem.  Suppose the probability of a correct solution is 0.1.  What is the probability that all three prizes will be awarded by the 25th submission?
```{r}

```








