---
title: "HW: Calculus of belief"
output: md_document
---

Consider two investment opportunities.  The first is a large water well project with a potential payoff of $100.  The second opportunity is an investment series of 10 small wells, each with a potential payoff of $10.  The probability that any well (large or small) returns a payoff is 0.1.

The functions below simulate the outcomes from the two investments.

```{r}
big_well <- function(N) 100*rbinom(N, 1, 0.1)
sma_well <- function(N)  10*rbinom(N, 10, 0.1)
```

For example, to simulate 10 outcomes, we can use the commands `big_well(10)` or `sma_well(10)`.

```{r}
big_well(10)
sma_well(10)
```

Thinking about probability as a long-run proportion, one can generate many replicates and then tabulate.

```{r}
big_well(1E6) |> 
  table() 

big_well(1E6) |>
  table() |>
  proportions()
```

1. Write code that generates the long-run proportions for the small well investment opportunity.

1. Create a figure which visualizes the long-run proportions.  (I know we have not covered this in-class; I'm asking you to search online for how you might do this.)

Please read chapter 3 of Understanding Uncertainty.

1. In section 3.9, the author introduces the notation $p(E|K)$. 
  1. What does $K$ represent?  
  1. Create an example where the probability of an event is different when $K$ is different.
1. If $E$ denotes an event, what does $E^c$ denote?
1. Consider a potentially unfair six-sided die.  Let $E$ denote the outcome of a die roll.  If $P(E$ is even $) = .6$, what must $P(E$ is odd $)$ be?  Why?

### Submission instructions

1.  Within your course repo, create a file called `HW5.md` with your answers formatted with markdown.
1. Sync your local and online repositories.  (Push the local repo to Github.com.)
1.  The deliverable should be your own work.  You may **discuss**
    concepts with classmates.

