---
output: md_document
---

# Error as a function of the number of replicates

```{r}
# Function which includes the replicate step
error_v4 <- function(r,n,p){
    d1 <- rbinom(r,n,p)
    phat <-d1/n
    ae <- abs(phat - p)
    re <- ae/p
    c(mean(ae), mean(re))
}

ns <- 2^(2:15)
ps <- c(0.01, 0.05, 0.1, 0.25, 0.5)
out <- expand.grid(N=ns, P=ps, AE = NA, RE = NA, KEEP.OUT.ATTRS = FALSE)

for(i in 1:nrow(out)){
  out[i, 3:4] <- error_v4(10000,n=out$N[i],p=out$P[i])
}

a1 <- array(out[,"AE"],dim=c(length(ns), length(ps)))

matplot(log2(a1), type = "b", pch = 16, axes=FALSE, ylab = "Absolute error", xlab = "Replicates")
axis(1,1:14,parse(text=paste0("2^",2:15)))
axis(2, axTicks(2), parse(text=paste0("2^",axTicks(2))), las = 1)
box()
grid()
```

Note that relationship between replicates and absolute error is linear when plotted on the log scale.

\[
\log_2(E) = \alpha + \beta \log_2(R)
\]

Looking at the plot, we see that the slope, $\beta$, is $-\frac{1}{2}$.  The formula can be reexpressed in the following ways:

\[
\begin{aligned}
\log_2(E) &= \alpha - \frac{1}{2} \log_2(R) \\
2^{\log_2(E)} &= 2^{\alpha - \frac{1}{2} \log_2(R)} \\
E &= 2^{\alpha}R^{-\frac{1}{2}} \\
 &= \frac{2^{\alpha}}{\sqrt{R}}
\end{aligned}
\]

This equation gives a way to understand the relationship between the number of replicates and the error. With it, we can answer the question: "If we want to reduce the error by half, how many additional replicates are needed?"

\[
\begin{aligned}
\frac{1}{2}E &= \frac{1}{2}\frac{2^{\alpha}}{\sqrt{R}} \\
&= \frac{2^{\alpha}}{2\sqrt{R}} \\
&= \frac{2^{\alpha}}{\sqrt{4R}} \\
\end{aligned}
\]

>**IMPORTANT POINT:** In order to reduce the error by half, the number of replicates must be **quadrupled**!

We can see this looking at specific values of $p$ and $n$.

## N = 100 & 400
```{r}
error_v4(100000, 100, 0.5)/2
error_v4(100000, 400, 0.5)
```

## N = 1000 & 4000
```{r}
error_v4(100000, 1000, 0.5)/2
error_v4(100000, 4000, 0.5)
```
