---
title: "Deliverable 1 Solutions"
output: md_document
date: "2024-04-03"
---

## Introduction

Below is code to help update your Deliverable 1 reports. I have not included the written portions of the blog post, but the code should help you create the necessary components to then explain.

## Provided Code

```{r, warning = FALSE}

single_spin <- function(){
  possible_outcomes <- c(rep("red",18), rep("black",18), rep("green",2))
  sample(possible_outcomes, 1)
}

martingale_wager <- function(
    previous_wager
  , previous_outcome
  , max_wager
  , current_budget
){
  if(previous_outcome == "red") return(1)
  min(2*previous_wager, max_wager, current_budget)
}

one_play <- function(previous_ledger_entry, max_wager){
  # Create a copy of the input object that will become the output object
  out <- previous_ledger_entry
  out[1, "game_index"] <- previous_ledger_entry[1, "game_index"] + 1
  out[1, "starting_budget"] <- previous_ledger_entry[1, "ending_budget"]
  out[1, "wager"] <- martingale_wager(
    previous_wager = previous_ledger_entry[1, "wager"]
    , previous_outcome = previous_ledger_entry[1, "outcome"]
    , max_wager = max_wager
    , current_budget = out[1, "starting_budget"]
  )
  out[1, "outcome"] <- single_spin()
  out[1, "ending_budget"] <- out[1, "starting_budget"] + 
    ifelse(out[1, "outcome"] == "red", +1, -1)*out[1, "wager"]
  return(out)
}

one_series <- function(
  max_games, starting_budget, winning_threshold, max_wager
){
  # Initialize ledger
  ledger <- data.frame(
      game_index = 0:max_games
    , starting_budget = NA_integer_
    , wager = NA_integer_
    , outcome = NA_character_
    , ending_budget = NA_integer_
  )
  ledger[1, "wager"] <- 1
  ledger[1, "outcome"] <- "red"
  ledger[1, "ending_budget"] <- starting_budget
  for(i in 2:nrow(ledger)){
    #browser()
    ledger[i,] <- one_play(ledger[i-1,], max_wager)
    if(stopping_rule(ledger[i,], winning_threshold)) break
  }
  # Return non-empty portion of ledger
  ledger[2:i, ]
}

stopping_rule <- function(
  ledger_entry
  , winning_threshold
){
  ending_budget <- ledger_entry[1, "ending_budget"]
  if(ending_budget <= 0) return(TRUE)
  if(ending_budget >= winning_threshold) return(TRUE)
  FALSE
}

profit <- function(ledger){
  n <- nrow(ledger)
  profit <- ledger[n, "ending_budget"] - ledger[1, "starting_budget"]
  return(profit)
}

require(magrittr)

svg(filename = "loser.svg", width=16, height =9)
par(cex.axis=2, cex.lab = 2, mar = c(8,8,2,2), bg = rgb(222, 235, 247, max = 255))
set.seed(1)
ledger <- one_series(200,200,300,500)
plot(ledger[,c(1,5)], type = "l", lwd = 5, xlab = "Game Index", ylab = "Budget")
dev.off()

svg(filename = "winner.svg", width=16, height =9)
par(cex.axis=2, cex.lab = 2, mar = c(8,8,2,2), bg = rgb(222, 235, 247, max = 255))
set.seed(2)
l2 <- one_series(200,200,300,500)
plot(l2[,c(1,5)], type = "l", lwd = 5, xlab = "Game Index", ylab = "Budget")
dev.off()

```

## Time Series

You can run with different seeds 1-3 times to get a sense of the variability of a player's budget trajectory. It's good to show that it can be either positive (winning) or negative (losing).

```{r}

# create a ledger using the default parameters
# default parameters come from the assignment instructions

set.seed(20302)
ledger <- one_series(1000, 200, 300, 100)

# showing what the ledger is
head(ledger)

# plot the ledger
plot(ledger$game_index, ledger$starting_budget, 
     xlab = 'Game Index', 
     ylab = 'Budget',
     main = "Player's Budget Over A Single Series",
     type = 'l')

```

## Average Earnings

We want to replicate the one_series() command many times, calculate the profit each time, and then take the mean after all of the replications.

```{r}

# we can do one_series 1000 times then take mean profit
# using the default parameters from the instructions

replicate(1000, one_series(1000, 200, 300, 100) |> profit()) |> mean()

```

## Changing Parameters Effect

You can repeat this process for the other parameters as well. This example is looking at starting budget. Just make sure that that all combinations of the parameters makes sense. For example, you don't want the starting budget to be greater than the winning threshold, otherwise the game stops right away.

```{r}

# ex. Starting Budget

# get a range of starting budgets
starting_budgets <- seq(100, 500, by = 100)

# create a vector to store the average earnings for each starting budget
# make it the length of the starting budgets we are looping over
ae <- numeric(length(starting_budgets))

# create a for loop, going through each starting budget value
for(i in 1:length(starting_budgets)){
  # replicate 1000 times with these settings, changing the starting budget each time
  # winning threshold is 100 more than the starting budget each time
  # could also just set winning threshold to be a large number that starting budget
  # never goes over
  # ae[i] stores the average earning in the ith place
  # starting_budgets[i] selects the ith starting budget to use for this loop
  ae[i] <- replicate(1000, one_series(1000, 
                                      starting_budgets[i], 
                                      starting_budgets[i] + 100, 
                                      100) |> profit()) |> mean()
}

# now plot the results
# starting budget on the x-axis, average earnings on the y-axis

plot(starting_budgets, ae, 
     main = 'Effect of Starting Budgets on Average Earnings',
     ylab = 'Average Earnings',
     xlab = 'Starting Budget',
     type = 'l')

```

## Number of Plays

This can be a plot as well across a certain parameter value. This is calculating with the default parameters to get a point estimate.

```{r}

# each row in the ledger is one play, we can get the number of plays using the number of rows

replicate(1000, one_series(1000, 200, 300, 100) |> nrow()) |> mean()

```