---
title: "HW 14 Solutions"
date: "2024-04-06"
output: pdf_document
---

# Setup Code

```{r}

# 1. define the values
k <- 4
p_A <- 0.55

# 2. Sequence of Length
s <- rbinom(2*k-1,1,p_A)

# 3. Determine if A or B won the series
# 1s will be team A winning

o <- ifelse(sum(s) >= 4, 'A', 'B')

# 4. Number of Games Played

# loop through the possible lengths of games
for(i in k:(2*k - 1)){
  # A is the number of games A won 
  A = sum(s[1:i])
  B = i - A
  
  # check if A or B won 4 games yet
  if((A == 4) | (B == 4)){
    # number of games is i
    n = i
    
    # leave the for loop since we know the number of games now
    break
  }
}

# 5. Put into a Function called world series

world_series <- function(k, p_A){
  
  # create the sequence of games
  s <- rbinom(2*k-1,1,p_A)
  
  # get the winner of the series
  o <- ifelse(sum(s) >= 4, 'A', 'B')
  
  # get the number of games
  for(i in k:(2*k - 1)){
    # A is the number of games A won 
    A = sum(s[1:i])
    B = i - A
    
    # check if A or B won 4 games yet
    if((A == 4) | (B == 4)){
        # number of games is i
        n = i
        
        # leave the for loop since we know the number of games now
        break
      }
  }
  
  # package as a vector
  return(c(o, n))
  
}

# 6. Create a function with inputs R, k, p_A, output an R x 2 with o in Col 1 and n in Col 2

replicate_world_series <- function(R, k, p_A){
  replicate(R, world_series(k, p_A)) |>
    t() |>
    matrix(ncol = 2, nrow = R)
}

```

# 1. Generate 5000 draws from the world series distribution (Use any k and $p_A$).

```{r}

R <- 5000
k <- 4
p_A <- 0.55

# set the seed for reproducibility
set.seed(20302)
output <- replicate_world_series(R, k, p_A)

# make output a Data Frame
output <- output |> 
  data.frame()

# df column names
colnames(output) <- c('Winner', 'NumGames')

# print first 5 rows of the output
head(output, 5)

```

# 2. With the draws, create the contingency table with number of games as the row variable and winner as the column variable. Include conditional probabilities.

```{r}
# load gmodels for cross table command
library(gmodels)

# create the cross table 
gmodels::CrossTable(output$NumGames, 
                    output$Winner, 
                    prop.chisq = FALSE)
```

# 3. Supposing that the world series ends in $2k - 1$ games, what is the probability that team A will win?

We can get this value from the table we produced. The probability we want is $P(\text{A} | \text{7 Games})$. From the cross table, this is the row conditional probability for 7 Games and A wins which is 0.549. This matches our value for $p_A$.
