---
output: md_document
---

# HW 3 Solutions

1. Working Directory

* What is the working directory?

>The working directory is a folder in which the program will look for input files and save output files, unless the location of the file is specified.  For example, the command `read.csv("survey.csv")` will look for the `survey.csv` file in the working directory.

* What are the best practices for setting and using the working directory?

>Write your code assuming that the working directory is the folder in which the code will be executed.

>Use **relative file paths**.  For example, use **..** syntax.

2. Create a string called `name` which stores your name.

```{r}
name <- "Thomas Stewart"
```

3. Run the `ls()` command to see the objects in the environment.

```{r}
ls()
```

4. Create vectors of length 5 with the following atoms.  Call them `v1`, `v2`, ..., `v4`.

* logicals
* integers
* doubles
* characters

```{r}
v1 <- c(TRUE, TRUE, FALSE, FALSE, TRUE)
v2 <- c(1L, 0L, 9L, 100000L, 4L)
v3 <- c(sqrt(2), pi, (1+sqrt(5))/2, 2.0, 5.1)
v4 <- c("This is", "a", "vector of", "character strings",".")
```

5. Place the vectors `v1` ... `v4` into a list called V.

```{r}
V <- list(v1, v2, v3, v4)
```

6. Use `str` to inspect `V`

```{r}
str(V)
```

7. Create a matrix of size 4 x 5 of numeric values.

```{r}
M <- matrix(rnorm(20), 4)
M
```

8. Create a sequence from 22 to 44 using the `seq` command.

```{r}
seq(22,44)
```

9. Create an array with `cbind`

```{r}
cbind(A = v3, B = 1:5)
```

10. Grab the 3rd element of the vector `v3`

```{r}
v3[3]
```

11. Use the following data frame, `d1`, and grab the 4th column.

```{r}
d1 <- read.csv("http://hbiostat.org/data/repo/2.20.Framingham.csv")
head(d1[,4]) # Show only first 6 elements
head(d1[[4]])
```

12. Use `d1` and grab the 5th row.

```{r}
d1[5,]
```

13. Use `d1` and grab the `bmi` column.

```{r}
# These are all equivalent
z1 <- d1[,"bmi"]
z2 <- d1$bmi
z3 <- d1[["bmi"]]
head(z1, 8) # First 8 entries
```

14. Run the following code chunk.  Explain what the pipe operator does.

```
require(dplyr)

# Old way
sum(log(sqrt(select(d1,age)),base=10))

# OG magrittr pipe
d1 %>% 
  pull(age)  %>% 
  sqrt  %>% 
  log(base=10)  %>% 
  sum

d1 %>% 
  lm(sbp ~ dbp, data = .)  # Use the dot if the piped objects needs to go to an input other than the first

d1 %>% 
  split(.$sex) %>%  # Can use the input multiple times
  lapply(function(x){x$age %>% mean})

# New base R pipe
d1 |> 
  pull(age) |> 
  sqrt() |> # Note the parentheses have to be used
  log(base=10) |> 
  sum()

d1 |>  
  lm(sbp ~ dbp, data = _) # Use `_` instead of `.` with new pipe
```

>The pipe operator allows the output of one command to be used as an input in a subsequent command.  In this way, the pipe creates a chain of commands to generate a final output.  Because the commands are written in the order in which they are executed, the code is easier to read.

15.  UVA phone numbers are often of the form (434) 924-????.  Modifying the code below, find the number of UVA phone numbers of the form (434) 924-???? which are prime.  (You may need to install the `primes` package.)

```{r}
prime <- function(x){
  ps <- primes::generate_primes(min = 2, max = ceiling(sqrt(x)))
  for(i in ps) if(x%%i == 0) return(FALSE)
  TRUE
}

k <- 0
for(i in 4349240000:4349249999){
    k <- k + prime(i)
}
k
```

16. Modify the code above to find the list UVA phone numbers of the form (434) 924-???? which are prime.

```{r}
out <- integer()
for(i in 4349240000:4349249999){
  if(prime(i)) out <- c(out,i)
}
head(out, 21) # Listing the first 21
```